Sampbias, a method to evaluate geographic sampling bias in species distribution data

### 2019-12-03 manuscript.rmd
Update manuscript.rmd

The rate parameter $l$ describes the strength of the bias effect. When  $l$ is large the expected probability of sampling occurrences decreases very quickly as you move away from a bias source (Fig. \ref{fig:distdi}). In contrast, when the $l$ parameter is small, the resulting exponential distribution effectively becomes more and more similar to a uniform distribution (Fig. \ref{fig:distdi}), indicating that increasing distance from e.g. a city does not affect the sampling probability. We treat $l$ as an unknown variable and estimate it using maximum likelihood from the data based on the probability density function described by $g(x,l)$. This essentially means finding the value of $l$ that best explains any discrepancies between the expected distribution ($f(x)$) and the observed occurrences.
[[*I wonder if this isn't in fact a posterior probability where f(x) is the prior and b(x,l) the likelihood or something like that*]]

@@

# Assumptions and future prospective
Two assumptions of *sampbias* are a equal sampling of occurrence records across the study area as null model and an exponential increase of the biasing effect with distance from the gazetteers. We considered both acceptable approximations for the purpose of the package, but future expansions of *sampbias* could relax these assumptions, for instance by allowing other distance decay functions, such as a gamma distribution, and by changing the sampling scheme of the background points. The first steps towards these goals are already implemented in the current version of *sampbias* with the option to limit background points to a convex hull around the dataset or limiting background points to terrestrial surface.

>>>>

The rate parameter $l$ describes the strength of the bias effect. When  $l$ is large the expected probability of sampling occurrences decreases very quickly as you move away from a bias source (Fig. \ref{fig:distdi}). In contrast, when the $l$ parameter is small, the resulting exponential distribution effectively becomes more and more similar to a uniform distribution (Fig. \ref{fig:distdi}), indicating that increasing distance from e.g. a city does not affect the sampling probability. We treat $l$ as an unknown variable and estimate it using maximum likelihood from the data based on the probability density function described by $g(x,l)$. This essentially means finding the value of $l$ that best explains any discrepancies between the expected distribution ($f(x)$) and the observed occurrences.
[[*I think this is in fact a posterior probability where f(x) is the empirical prior and b(x,l) the likelihood and we do a maximum a posteriori optimization*]]

@@

# Assumptions and future prospective
Two assumptions of *sampbias* are a equal sampling of occurrence records across the study area as null model and an exponential increase of the biasing effect with distance from the gazetteers. We considered both acceptable approximations for the purpose of the package, but future expansions of *sampbias* could relax these assumptions, for instance by allowing other distance decay functions, such as gamma or Weibull distributions, and by changing the sampling scheme of the background points. The first steps towards these goals are already implemented in the current version of *sampbias* with the option to limit background points to a convex hull around the dataset or limiting background points to terrestrial surface. [[*I don't understand this part*]]
